---
title: "CMNH Data Processing"
output:
  html_document:
    df_print: paged
---

# Overview

This script cleans the data provided by the Cleaveland Museum of Natural History, herbarium specimens for *F. alnus* and the group of associated species.

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "/Users/maiellolammens/Dropbox/Projects/FRAL-herbarium/")
```

# Load necessary packages

```{r}
library(readxl)
library(dplyr)
```


# Read in the data in Excel format

```{r}
fral <- read_xlsx("data/CMNH/CMNH glossy buckthorn Data Revised.xlsx", sheet = 1)
alin2 <- read_xlsx("data/CMNH/CMNH Alnus incana Data.xlsx", sheet = 1)
alse2 <- read_xlsx("data/CMNH/CMNH Alnus serrulata Data.xlsx", sheet = 1)
rhal <- read_xlsx("data/CMNH/CMNH Rhamnus alnifolia Data.xlsx", sheet = 1)
fram2 <- read_xlsx("data/CMNH/CMNH Fraxinus americana Data.xlsx", sheet = 1)
havi4 <- read_xlsx("data/CMNH/CMNH Hamamelis virginiana Data.xlsx", sheet = 1)
```

How many counties and towns are represented in the FRAL data set?

```{r}
length(unique(fral$County))
length(unique(fral$`City/Twp.`))
```

```{r}
sort(unique(fral$`City/Twp.`))
```


# Combine all of the data sets and format

Combine the data sets with a consistent set of columns.

```{r}
cmnh_data <- bind_rows(select(fral, Genus, Species, `Herbarium No.`, Day, Month, Year, Collector,
                              Locality, Area, `City/Twp.`, County, State, Comments, Description),
                       select(alin2, Genus, Species, `Herbarium No.`, Day, Month, Year, Collector,
                              Locality, Area, `City/Twp.`, County, State, Comments, Description),
                       select(alse2, Genus, Species, `Herbarium No.`, Day, Month, Year, Collector,
                              Locality, Area, `City/Twp.`, County, State, Comments, Description),
                       select(fram2, Genus, Species, `Herbarium No.`, Day, Month, Year, Collector,
                              Locality, Area, `City/Twp.`, County, State, Comments, Description),
                       select(havi4, Genus, Species, `Herbarium No.`, Day, Month, Year, Collector,
                              Locality, Area, `City/Twp.`, County, State, Comments, Description),
                       select(rhal, Genus, Species, `Herbarium No.`, Day, Month, Year, Collector,
                              Locality, Area, `City/Twp.`, County, State, Comments, Description)
                       )
```

Convert `Day` and `Year` columns to numeric valeus.

```{r}
cmnh_data$Day <- as.numeric(cmnh_data$Day)
cmnh_data$Year <- as.numeric(cmnh_data$Year)
```

Remove entries with `Year == 0`.

```{r}
cmnh_data <-
  cmnh_data %>%
  filter(Year != 0)
```

# Add `County` data

I looked up county locations for those that were missing this field using the Township and State information.
All look-ups were done in Google Maps.

```{r}
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "033498")] <- "Hudson"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "040830")] <- "Grafton"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "063821")] <- "Cobb"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "032557")] <- "Franklin"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "032558")] <- "Grafton"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "041369")] <- "Lee"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "041372")] <- "Grafton"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "041375")] <- "Hampshire"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "062232")] <- "Walton"
cmnh_data$County[which(cmnh_data$`Herbarium No.` == "062234")] <- "Cook"
```

Drop "Co." part of any of the county entries.

```{r}
cmnh_data$County <- sub(pattern = " Co\\.", replacement = "", x = cmnh_data$County)
```

Remove any records that do not have County noted at this point

```{r}
cmnh_data <- filter(cmnh_data, !is.na(County))
```


Look at state abreviations for remaining data

```{r}
table(cmnh_data$State)
```

Fix the "Oh" values

```{r}
cmnh_data$State[which(cmnh_data$State == "Oh")] <- "OH"
```



# Georeferencing of records at the County Level

I'm going to georeference each of these locations at the county level, using
data from the US Census Bureau. Specifically I am using data from the 
**US Census Gazetteer**
acquired from this website: 
[http://www.census.gov/geo/maps-data/data/gazetteer2010.html](http://www.census.gov/geo/maps-data/data/gazetteer2010.html) 

I downloaded the 'Counties' data set as a *.txt file. I opened it in Excel and saved
it as a *.csv file. I also copied the metadata information from this website and saved
it as a separate *.txt file. All of the Gazetteer files are currently stored in my 
'gis_layers' directory in my Google Drive.

```{r}
# Read in the counties *.csv file
Counties <- read.csv('~/Google Drive//gis_layers/Gaz_counties_national.csv')
# Remove the word 'County' in the 'NAME' column
Counties$NAME <- sub(pattern=' County',replacement='',Counties$NAME)


cmnh_data <- merge(cmnh_data,Counties,
                   by.x=c('State','County'),by.y=c('USPS','NAME'))
# How many records remaining?
nrow(cmnh_data) # 433 - only lost 5 records, likely Canadian records

# Calculate a coordinate uncertainty value for these records
# Base the uncertainty on the radius of a circle of the area
# of the count. This is a rough approximation.
cmnh_data$CoordUncertainty <- sqrt(cmnh_data$ALAND/pi)


```


# Put CMNH data into the same formate as the other data sets

```{r}
## Make a clean SpeciesName column
cmnh_data$SpeciesName <- paste(cmnh_data$Genus, sub(pattern = "[[:blank:]].*", replacement = "", cmnh_data$Species))

## Add an herbarium code column
cmnh_data$HerbariumCode <- "CLM"

## Add a source column
cmnh_data$Source <- "From Curator"

## Convert from month chr to num
cmnh_data$Month <- match(cmnh_data$Month, month.name)

## Make a country column
cmnh_data$Country <- "United States"

## Make a Analysis notes column
cmnh_data$AnalysisNotes <- ""

## Make a Remove column
cmnh_data$Remove <- 0

## Make a County only column
cmnh_data$CountyLevelOnly <- 1

cmnh_data_final <- cmnh_data %>%
  select(SpeciesName = SpeciesName, 
         HerbariumCode = HerbariumCode,
         Source = Source,
         Collector = Collector,
         RecordNumber = `Herbarium No.`,
         CollectionDay = Day,
         CollectionMonth = Month,
         CollectionYear = Year,
         Country = Country,
         Admin1 = State,
         Admin2 = County,
         Admin3 = `City/Twp.`,
         Locality = Locality, 
         Longitude = INTPTLONG,
         Latitude = INTPTLAT,
         CoordUncertainty = CoordUncertainty, 
         LocationNotes = Description,
         EcologyNotes = Comments,
         AnalysisNotes = AnalysisNotes, 
         Remove = Remove,
         CountyLevelOnly = CountyLevelOnly)

write.csv(cmnh_data_final, "data/CMNH_Data_Cleaned.csv", row.names = FALSE)
```

